WDIAS : A Microservices-Based Weather Data Integration and Assimilation System\\

### 2020-12-30 lit/sec_3-other-systems.tex
Chap 5: Added discussion paragraph regarding existing systems vs WDIAS

\acrfull{msm} framework develops an open system architecture which consists of open-data concept to access heterogeneous data sources via data agents, and open-model concept to allow individual model handling via model agents. Even through these features are provided by other systems such as \acrshort{fews}, they do not provide decentralized model execution and data integration. Specifically, this framework provides a graphical user interface for workflow creation similar to \acrshort{lead}, rather than providing a configuration-based workflow system as in \acrshort{fews}. In our research, we are not mainly focusing on implementing a workflow engine, we will not look deep into such features here.
Then \acrshort{msm} tries to effectively address the drawbacks of the existing modeling systems by providing features such as an independent open-source and graphical-based workflow engine, facilitating  users to obtain data directly from popular sources, decentralized model administration by modelers, and enables models to be freely connected with each other in the \acrshort{msm} framework.

@@

\cref{fi:msm_architecture} shows the \acrshort{msm} system framework from a top-level design point of view, and it is mainly composed of the \acrshort{msm} core which is an interface with the workflow engine, data agents, and model agents. Thus, the \acrshort{msm} core can become the bottleneck of the system, and a single point of failure in critical situations. The \acrshort{msm} core also controls all other components and by using corresponding model agents and data agents allows it to connect with external models and remote data sources. The \acrshort{msm} core is connected to the workflow engine and provides the end-user with all workflow control capabilities, and it contains data persistence services that can be used by all \acrshort{msm} components. Another drawback is, implementing new data agents and model agents are programming language-dependent. In order to implement Model Agents and Data agents, users need to inherit \acrshort{msm} classes in the provided language. Then users need to upload the code into a folder (not a containerized mechanism, or continuous service deployments).

The open-data architecture of \acrshort{msm} is using to access multiple and heterogeneous external data sources. The users are able to integrate data from any source via data agents, but the data agents are depending on inheriting a generic class to do so. Which means there is a language dependency. The open model architecture allows users to independently develop and deploy their own model via model agents. Here also users have to use a generic agent provided by the \acrshort{msm} framework, but the models can be implemented independently. Users can deploy new model agents by copy pasting the source code to a specific folder, but it is possible to run the models which can be written in C, C++, FORTRAN, Python, Java, etc. During the model execution, it follows a flow similar to the Delft-FEW model adapter with a prepared input file, execute the model, and processing output files. The \acrshort{msm} framework has a data management layer, which stores the data as an in memory and local data caching, featuring a non-relational database scheme that supports high throughput. It stores particular timeseries data as msmDataSets, and creates a unique identifier for each msmDataSet. 
The msmDataSets storage framework contains the time-steps. Each time-step contains a reference to the msmDataSet it belongs to, and a Document (the data oriented structure) with a 2D image.

In this paper, it presents a use case of using the \acrshort{msm} system, and does not provide detailed evaluation regarding a performance load testing. In this use case, it demonstrates the usefulness and effectiveness of the framework in various aspects of scientific modeling activities. The msm has only been fully tested on Windows 10 operating system, which means the framework has some operating system dependency. Further, It is not very clear how is it parallel running data agents, model agents, and the \acrshort{msm} core are. Even though it mentioned it follows the decentralized agents approach, based on the available data we can conclude it also has a monolithic distributed architecture.

>>>>

\gc{\acrfull{msm} framework develops an open system architecture which consists of open-data concept to access heterogeneous data sources via data agents, and open-model concept to allow individual model handling via model agents. Even through these features are provided by other systems such as \acrshort{fews}, they do not provide decentralized model execution and data integration. Specifically, this framework provides a graphical user interface for workflow creation similar to \acrshort{lead}, rather than providing a configuration-based workflow system as in \acrshort{fews}. In our research, we are not mainly focusing on implementing a workflow engine, we will not look deep into such features here.
Then \acrshort{msm} tries to effectively address the drawbacks of the existing modeling systems by providing features such as an independent open-source and graphical-based workflow engine, facilitating  users to obtain data directly from popular sources, decentralized model administration by modelers, and enables models to be freely connected with each other in the \acrshort{msm} framework.}

@@

\gc{\cref{fi:msm_architecture} shows the \acrshort{msm} system framework from a top-level design point of view, and it is mainly composed of the \acrshort{msm} core which is an interface with the workflow engine, data agents, and model agents. Thus, the \acrshort{msm} core can become the bottleneck of the system, and a single point of failure in critical situations. The \acrshort{msm} core also controls all other components and by using corresponding model agents and data agents allows it to connect with external models and remote data sources. The \acrshort{msm} core is connected to the workflow engine and provides the end-user with all workflow control capabilities, and it contains data persistence services that can be used by all \acrshort{msm} components. Another drawback is, implementing new data agents and model agents are programming language-dependent. In order to implement Model Agents and Data agents, users need to inherit \acrshort{msm} classes in the provided language. Then users need to upload the code into a folder (not a containerized mechanism, or continuous service deployments).}

\gc{The open-data architecture of \acrshort{msm} is using to access multiple and heterogeneous external data sources. The users are able to integrate data from any source via data agents, but the data agents are depending on inheriting a generic class to do so. Which means there is a language dependency. The open model architecture allows users to independently develop and deploy their own model via model agents. Here also users have to use a generic agent provided by the \acrshort{msm} framework, but the models can be implemented independently. Users can deploy new model agents by copy pasting the source code to a specific folder, but it is possible to run the models which can be written in C, C++, FORTRAN, Python, Java, etc. During the model execution, it follows a flow similar to the Delft-FEW model adapter with a prepared input file, execute the model, and processing output files. The \acrshort{msm} framework has a data management layer, which stores the data as an in memory and local data caching, featuring a non-relational database scheme that supports high throughput. It stores particular timeseries data as msmDataSets, and creates a unique identifier for each msmDataSet. 
The msmDataSets storage framework contains the time-steps. Each time-step contains a reference to the msmDataSet it belongs to, and a Document (the data oriented structure) with a 2D image.}

\gc{In this paper, it presents a use case of using the \acrshort{msm} system, and does not provide detailed evaluation regarding a performance load testing. In this use case, it demonstrates the usefulness and effectiveness of the framework in various aspects of scientific modeling activities. The msm has only been fully tested on Windows 10 operating system, which means the framework has some operating system dependency. Further, It is not very clear how is it parallel running data agents, model agents, and the \acrshort{msm} core are. Even though it mentioned it follows the decentralized agents approach, based on the available data we can conclude it also has a monolithic distributed architecture.}

### 2020-12-30 method/sec_5-data_preprocess.tex
Chap 5: Added discussion paragraph regarding existing systems vs WDIAS

The data preprocess functionalities support simple generic mathematical function models. Each extension is considered as a mathematical function that can take \texttt{p} input timeseries variables and output \texttt{n} timeseries variables after applying the given function. Further, the system supports to provide a bind constant as configurations to the extension at the time of preprocessing. Thus, users can change the behavior of an existing function by providing different bind constants at the time of creating new triggers for an extension and produce different behaviors.

\subsection{Interpolation}
The interpolation module generates data at desired locations or desired data points using a serial or spatial interpolation technique. It is used to fill gaps in on-line measured data and to derive spatially distributed data for weather timeseries, such as precipitation and temperature, based on information available at nearby locations.

\subsubsection{Serial Interpolation}
During serial interpolation, we use interpolation to fill any gaps in a timeseries. The interpolation module will only consider a single timeseries, which is itself in filling these gaps.

@@

\subsubsection{Spatial Interpolation}
During spatial interpolation, we use interpolation to either fill gaps in timeseries or to create a new timeseries for a location using data from other (spatially distributed) locations. Spatial interpolation can also be applied for sampling scalar timeseries from grid timeseries, for re-sampling grids, or for creating grids from timeseries data.
\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{method/data_preprocess/spatial_interpolation.pdf}
    \caption{Spatial interpolation.}
    \label{fi:spatial_interpolation}
\end{figure}

\subsection{Transformation}
Simple arithmetic manipulation of time interval transformation and shifting the series in time specific hydro-meteorological transformation such as stage discharge relationships.
\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{method/data_preprocess/transformation.pdf}
    \caption{Time interval aggregration.}
    \label{fi:transformation}
\end{figure}

\subsection{Validation}
Checks for counting reliable, doubtful, unreliable, and missing values.
\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{method/data_preprocess/validation.pdf}
    \caption{Timeseries data validation.}
    \label{fi:validation}
\end{figure}

>>>>

The data preprocess functionalities support simple generic mathematical function models. Each extension is considered as a mathematical function that can take \texttt{p} input timeseries variables and output \texttt{n} timeseries variables after applying the given function. Further, the system supports to provide a bind constant as configurations to the extension at the time of preprocessing. Thus, users can change the behavior of an existing function by providing different bind constants at the time of creating new triggers for an extension and produce different behaviors. \gkc{Next, we explain how can we derive interpolation, transformation, and validation by using the generic mathematical function model that we described above.}

\subsection{Interpolation}
\gkc{The interpolation extension generates data at desired locations or desired data points using a serial or spatial interpolation technique. It is used to fill gaps in linear measured data and to derive spatially distributed data for weather timeseries, such as precipitation and temperature, based on information available at nearby locations. Here, we will discuss two types of interpolation such as serial interpolation and spatial interpolation below.}

\subsubsection{Serial Interpolation}
\gkc{During serial interpolation, we use interpolation to fill any gaps in a linear timeseries. The interpolation module will only consider a single timeseries, which is itself filling the missing data points. As an example, there can be timeseries which has the data points of temperature which are collected using a weather station. There can be missing data in this timeseries, and we can use a simple linear interpolation algorithm to fill the missing values based on the existing data points.}

@@

\subsubsection{Spatial Interpolation}
\gkc{During spatial interpolation, we use interpolation to either fill gaps in timeseries or to create a new timeseries for a location using data from other (spatially distributed) locations' timeseries. Spatial interpolation can also be applied for sampling scalar timeseries from grid timeseries, for re-sampling grids, or for creating grids from timeseries data. As an example, let’s assume that there are precipitation timeseries data available for the location A and location B as shown in Figure x. Then we want to calculate the mean precipitation of nearby location using these two timeseries. We can use the spatial interpolation to get the mean precipitation of nearby location.}
\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{method/data_preprocess/spatial_interpolation.pdf}
    \caption{Spatial interpolation.}
    \label{fi:spatial_interpolation}
\end{figure}

\subsection{Transformation}
\gkc{Simple arithmetic manipulation of time interval transformation and shifting the series in time-specific hydro-meteorological transformation such as stage-discharge relationships. As an example, there can be a timeseries of precipitation with 1-minute sample data. There will be a model requirement that needs hourly data. So, we need to do accumulation data manipulation to create a timeseries of hourly data using 1-minute sample data.}
\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{method/data_preprocess/transformation.pdf}
    \caption{Time interval aggregation.}
    \label{fi:transformation}
\end{figure}

\subsection{Validation}
\gkc{Checks for counting reliable, doubtful, unreliable, and missing values. The weather stations can produce invalid observations due to technical errors in the sensors of the devices. If we feed those values directly to the models, the models will produce inaccurate forecasting. To avoid that, we need to modify or flag those values before feeding them to the numerical models. We can use validation extensions to modify or flag invalid data. As an example, a weather station can record temperature above 70 celsius due to sensor error or calibration issues. But in real-world, it is not possible to have such a temperature in nature. Thus we need to flag such temperature issues and need to repair the weather station as soon as possible.}
\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{method/data_preprocess/validation.pdf}
    \caption{Timeseries data validation.}
    \label{fi:validation}
\end{figure}

### 2020-12-30 method/sec_6-query.tex
Chap 5: Added discussion paragraph regarding existing systems vs WDIAS

The system provides timeseries metadata query endpoints to search for existing timeseries metadata within the system using primary key attributes.
Also, it provides a set of geo-query endpoints, and users can perform geo-search queries with sending geo-information in geoJson format. The following geo-queries are supported by the system, and a detailed description is presented in Appendix B.
\begin{itemize}
    \item Query locations within area
    \item Query parameters in locations
    \item Query timeseries by location
    \item Query timeseries by locations
    \item Query timeseries by Parameter
    \item Query timeseries in area
    \item Query timeseries in area by parameter
    \item Query all timeseries
\end{itemize}

>>>>

The system provides timeseries metadata query endpoints to search for existing timeseries metadata within the system using primary key attributes mentioned in the \cref{subse:timeseries_key_attributes}.
Also, it provides a set of geo-query endpoints, and users can perform geo-search queries with sending geo-information in geoJson format. The following geo-queries are supported by the system, and a detailed description is presented in Appendix B.
\begin{itemize}
    \item Search for available locations within the provided geoJson area in the query.
    \item Search for available parameters in provided locations in the query.
    \item Search for available timeseries for a provided location in the query.
    \item Search for available timeseries for the given locations in the query.
    \item Search for available timeseries for given parameter. E.g., get available waterlevel timeseries in the system.
    \item Search for available timeseries within the provided geoJson area in the query.
    \item Search for available timeseries within the provided geoJson area by parameter. E.g., get available waterlevel timeseries within a given area.
    \item Query for all timeseries in the system.
\end{itemize}

### 2020-12-30 summary/sec_1-conclusions.tex
Upgrade Future work
Chap 5: Added discussion paragraph regarding existing systems vs WDIAS

To enhance the accuracy of the weather forecast, we need to provide reliable and timely weather data for NWMs. Also, most of the existing weather data integration systems are based on basic client-server architectures or monolithic distributed system architectures or developed based on the best technology available at the time of developing those systems. However, most of the concepts are outdated with the technology advancement, and we cannot get the full advantage of modern computational models such as containerized applications and using the cloud computing to gain benefits of scalability, high availability, and low-cost operations. Even though there are many weather data integration systems, most of them are proprietary or closed source. Thus, for an island like Sri Lanka that has different kinds of weather seasons yearly, those software needs to be highly customized

@@

First, the architecture of \acrshort{wdias} starts with \acrshort{soa} using an \acrfull{esb} and then moved to actor model-based architecture. After comparing the disadvantages of using such architectures with the system requirements, we came up with modern microservice architecture. One of attractive feature of microservice architecture is independent nature of microservices enables to design a highly scalable modular system which can use to handle large volumes of data using a large resource pool. Also, we used many of the concepts of modern microservice architecture to create stateless, failover, redundant microservice to achieve such capabilities. \acrshort{wdias} implemented the microservice architecture on top of the \acrshort{k8s}, which is a container orchestration platform. We mapped each microservice of \acrshort{wdias} to a pod in the \acrshort{k8s} cluster. Also, each microservice exposed endpoints to the users which are only focus on supporting a specific operation or functionality which follows the concept of smart endpoints dumb pipes in \cref{subse:dumb_pipes}. For importing different format data into the system, we added microservice for each data format that known as import microservices. Then, these import microservices store the data via an appropriate database adapter microservice after converting to compatible data format. For handling multiple data formats efficiently, we introduced a database structure in WDIAS. Then, users can fetch data in the required format via using export microservices. Further, we enabled integrating extensional microservices which provides the data preprocessing capabilities, and it allow users to integrate new features as new extension microservices. We used geo-index to support geo-based search queries and metadata search queries. Also, the system provides asynchronous request handling based on the request size with maintaining status per request.

>>>>

To enhance the accuracy of the weather forecast, we need to provide reliable and timely weather data for NWMs. Also, most of the existing weather data integration systems are based on basic client-server architectures or monolithic distributed system architectures or developed based on the best technology available at the time of developing those systems. However, most of the concepts are outdated with the technology advancement, and we cannot get the full advantage of modern computational models such as containerized applications and using the cloud computing to gain benefits of scalability, high availability, and low-cost operations. Even though there are many weather data integration systems, most of them are proprietary or closed source. Thus, for an island like Sri Lanka that has different kinds of weather seasons yearly, those software needs to be highly customize.

@@

First, the architecture of \acrshort{wdias} starts with \acrshort{soa} using an \acrfull{esb} and then moved to actor model-based architecture. After comparing the disadvantages of using such architectures with the system requirements, we came up with modern microservice architecture. One of attractive feature of microservice architecture is independent nature of microservices enables to design a highly scalable modular system which can use to handle large volumes of data using a large resource pool. Also, we used many of the concepts of modern microservice architecture to create stateless, failover, redundant microservice to achieve such capabilities. \acrshort{wdias} implemented the microservice architecture on top of the \acrshort{k8s}, which is a container orchestration platform. We mapped each microservice of \acrshort{wdias} to a pod in the \acrshort{k8s} cluster. Also, each microservice exposed endpoints to the users which are only focus on supporting a specific operation or functionality which follows the concept of smart endpoints dumb pipes. For importing different format data into the system, we added microservice for each data format that known as import microservices. Then, these import microservices store the data via an appropriate database adapter microservice after converting to compatible data format. For handling multiple data formats efficiently, we introduced a database structure in WDIAS. Then, users can fetch data in the required format via using export microservices. Further, we enabled integrating extensional microservices which provides the data preprocessing capabilities, and it allow users to integrate new features as new extension microservices. We used geo-index to support geo-based search queries and metadata search queries. Also, the system provides asynchronous request handling based on the request size with maintaining status per request.

\gc{When compared to other existing weather data integration and assimilation systems, the WDIAS has the advantage of taking full advantage of modern cloud computing technologies to implement a highly scalable and high-available system with low-cost. Since it uses containerized applications to implement the microservices, it allows independent service deployments. Each microservice can scale independently without affecting other microservices, and capable of using suitable technology stack for the task. The \acrshort{fews}, \acrshort{madis} and \acrshort{\acrshort{msm}} have platform dependencies, and some of the deployments are done with copy-pasting source codes. The \acrshort{fews} and \acrshort{msm} provide centralized data storage mechanism which can become a bottleneck while processing a huge set of data in critical situations. The \acrshort{madis} and \acrshort{dias} also support data storage array to store data. The WDIAS and \acrshort{lead} can support a pool of resources based on the requirement.
Similar to other systems, WDIAS provides the data preprocessing capabilities, and users can integrate new data preprocessing capabilities by implementing as independent containerized microservice applications. Even though \acrshort{fews}, \acrshort{lead} and \acrshort{msm} provide workflow engine support and model execution support, these can be achieved using an external workflow management system by fulfilling the data requirements via WDIAS system. Further, WDIAS is an extendable open framework such that similar capabilities can be integrated as a new extension module system. Compared to other existing systems, the \acrshort{lead} is more scalable which is implemented using \acrshort{soa} architecture as a result of contributions of lots of research from multiple universities. But with the help of modern container orchestration frameworks such as Kubernetes and cloud computing technologies, those are able to provide high scalability and high availability as inherent capabilities. And those are mature enough to automatically provide fault-tolerance \cite{Scolati2019ADevices}. By implementing the deployment units as containerized applications based on the modern architecture patterns allow WDIAS to implement a highly scalable and high available system without heavy effort on those features.}
